---
title: "Introduction to RICdata"
author: "Maria Dermit"
date: "`r BiocStyle::doc_date()`"
package: "`r pkg_ver('RICdata')`"
output: 
   BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Introduction to RICdata}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
  %\VignetteEncoding{UTF-8}
bibliography: ["RICdata.bib"]
---

```{r required packages, echo = FALSE, warning=FALSE, results="hide"}
suppressPackageStartupMessages({
    library("BiocStyle")
    library("RICdata")
    library(magrittr)
    library(tidyverse)
})
```


\pagebreak

# Introduction

`RICdata` is a data package containing the data to analyse RNA interaction 
capture from the manuscript titled: Global analysis of RNA-binding protein 
dynamics by comparative and enhanced RNA interactome capture. This paper was 
published in Nature Protocols 16,27–60(2021).Peptide information extracted from
_PRIDE_ [PXD009789](https://www.ebi.ac.uk/pride/archive/projects/PXD009789).A 
total of 9 oligo DT capture and total cell lysate samples originated from mass 
spectrometry proteomics SILAC quantitative experiments [@Garcia-Moreno:2019].
The package also contains a reduced version of data contained in [ProtFeatures](https://www.huber.embl.de/users/befische/RBDmap/Rpackage/RBDmapHeLa.pdf)
[@Castello:2016]. This is  called `miniProtFeatures` and contains the sequence 
information of the proteins. 
miniProtFeature is a list with the following objects:

- _ProtSeq_  a AAStringSet object of length 69025 with protein sequences.
- _GeneName_ a character vector with ENSEMBL gene id.
- _Symbol_ a named character vector with ENSEMBL gene symbols.
GO annotation provided [here](https://www.huber.embl.de/users/befische/mRNAinteractome/Rpackage/index.html)is included as `ENSG2category` object. 
Index maps for all amino acids 4-mers to proteins is provided as `Index` object, 
and is used by the function `mapPeptides` included in RIC` package to reverse
peptides on a protein sequence database. All these data are required to run 
functions in `RIC` package to analyse RNA interaction capture data.

# Session Info

```{r sessioninfo}
sessionInfo()
```

