---
title: "Introduction to RICdata"
author: "Maria Dermit"
date: "`r BiocStyle::doc_date()`"
package: "`r pkg_ver('RICdata')`"
output: 
  BiocStyle::html_document:
    self_contained: yes
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: show
vignette: >
  %\VignetteIndexEntry{Introduction to RICdata}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8} 
bibliography: ["RICdata.bib"]
---

```{r required packages, echo = FALSE, warning=FALSE, results="hide"}
suppressPackageStartupMessages({
    library("BiocStyle")
    library("RICdata")
    library(QFeatures)
    library(magrittr)
    library(tidyverse)
})
```

# Introduction

`RICdata` is a data package containing the data to analyse RNA interaction 
capture from the manuscript titled: Global analysis of RNA-binding protein 
dynamics by comparative and enhanced RNA interactome capture. This paper was 
published in Nature Protocols 16,27–60(2021).

# Proteomics data overview

Peptide information extracted from
_PRIDE_ [PXD009789](https://www.ebi.ac.uk/pride/archive/projects/PXD009789).A 
total of 9 oligo DT capture and total cell lysate samples originated from mass 
spectrometry proteomics SILAC quantitative experiments [@Garcia-Moreno:2019].

```{r}
# Path to tabular data
WCLpeptidesfilepath<- system.file("extdata","WCL_peptides.txt", package = "RICdata")
RICpeptidesfilepath<- system.file( "extdata", "RIC_peptides.txt", package = "RICdata")

data("WCLpeptides.raw")
dim(WCLpeptides.raw)
data("RICpeptides.raw")
dim(RICpeptides.raw)
```

Indices of the columns to be used as expression values are as follow:
```{r}
j <- str_which(colnames(WCLpeptides.raw),str_c(c("Intensity.((\\D)).18_M_4",
                                             "Intensity.((\\D)).4_18_M",
                                             "Intensity.((\\D)).M_4_18"),
                                              collapse="|"))
colnames(WCLpeptides.raw)[j]

i <- str_which(colnames(RICpeptides.raw),str_c("Intensity.[H|M|L].",
                                               collapse="|"))
colnames(RICpeptides.raw)[i]
```

# Annotation data overview

`RICdata` package also contains a reduced version of data contained in [ProtFeatures](https://www.huber.embl.de/users/befische/RBDmap/Rpackage/RBDmapHeLa.pdf)
[@Castello:2016]. This object is called `miniProtFeatures` and contains proteins 
sequence information. 
miniProtFeature is a list with the following objects:

- _ProtSeq_  a AAStringSet object of length 69025 with protein sequences.
- _GeneName_ a character vector with ENSEMBL gene id.
- _Symbol_ a named character vector with ENSEMBL gene symbols.

```{r}
data(miniProtFeatures)
head(miniProtFeatures$ProtSeq)
head(miniProtFeatures$GeneName)
head(miniProtFeatures$Symbol)
```

GO annotation provided in [mRNAinteractome](https://www.huber.embl.de/users/befische/mRNAinteractome/Rpackage/index.html) is included and called `ENSG2category`. 

```{r}
data(ENSG2category)
head(ENSG2category)
```

Index maps for all amino acids 4-mers to proteins is provided as `Index` object, 
and is used by the function `mapPeptides` included in RIC` package to reverse
peptides on a protein sequence database.

```{r}
data(Index)
head(Index$AAAA)
```

All these data are required to run functions in `RIC` package in order to
analyse RNA interaction capture data.


# Session Info

```{r sessioninfo}
sessionInfo()
```

